### 2014-10-07 paper/include/background.tex
workwork

To refresh one rank, the simplest approach is to refresh all banks in parallel row-by-row at the start of the 64ms interval. This approach is called Burst refresh and has some downsides. The approach blocks normal accesses for a longer period and form a peak of power consumption during the burst time. A better alternative is to distribute the refreshes across the interval, resulting in a lower average latency for the processor and a more stable power consumption.

The implementation of a DRAM refresh cycle can be made in two different ways, to either affect one row or many. The first is called RAS-only refresh (ROR) where refresh cycle refreshes one targeted row. The address for the targeted row has to be put on the bus and it is the MC that makes sure that a row are refreshed within the 64ms interval. The second is called CAS before RAS refresh (CBR). In this scheme the memory module has an internal address counter that displays which row that shall be refreshed. On a refresh cycle, the row stated by the internal address counter is refreshed and the counter is then incremented. The advantage with this method is that the address is never put on the bus, which gives a lower power consumption and less bus usage than ROR.

@@

\todo[inline]{Drafty stuff below}

In the next years it is projected that ... blablabla [source] so it is important and will be more important in the future to lower the total DRAM power consumption and an relatively easy way to do this is to optimizing the DRAM refresh. Next. we present a bunch of techniques that target DRAM refresh optimization.

>>>>

To refresh one rank, the simplest approach is to refresh all banks in parallel row-by-row at the start of the 64 ms interval. This approach is called Burst refresh and has some downsides. The approach blocks normal accesses for a longer period and form a peak of power consumption during the burst time. A better alternative is to distribute the refreshes across the interval, resulting in a lower average latency for the processor and a more stable power consumption.

The implementation of a DRAM refresh cycle can be made in two different ways, to either affect one row or many. The first is called RAS-only refresh (ROR) where refresh cycle refreshes one targeted row. The address for the targeted row has to be put on the bus and it is the MC that makes sure that a row are refreshed within the 64 ms interval. The second is called CAS before RAS refresh (CBR). In this scheme the memory module has an internal address counter that displays which row that shall be refreshed. On a refresh cycle, the row stated by the internal address counter is refreshed and the counter is then incremented. The advantage with this method is that the address is never put on the bus, which gives a lower power consumption and less bus usage than ROR.

@@

The conclusion drawn from \reffig{fig:dram_data_proj} is that it will be increasingly important to mitigate the throughput loss and DRAM power consumption of future systems. One way to decrease these negative trends of DRAM scaling is to optimize the DRAM refresh.

### 2014-10-07 paper/include/smart_refresh.tex
workwork

The basic concept behind our scheme is that a memory line that has been recently read out or written to does not need to be refreshed again by the periodic refresh mechanism.

Uses a time-out counter for each memory line. The counter is set on a read/write/refresh and dicards refreshes until the timer hits zero. The counter is associated to a bank/line pair. The counters are stored in the memory controller. One counter consists of 2 SRAM bits. More bits give higher granularity and thus also better performance because the refresh is made less often.

@@

Implemented so only one line per each logical segments are deremented at the same time, which leads to that the refreshes are evenly distributed over time. (I.e. no burst refresh.)

@@

Employed its solution to 2 GB ram. What about todays sizes? How effective is the technique now? - Only works effective if large parts the memory is acccessed, i.e. large part of the memory has to be valid to begin with. Is it so today?

>>>>

M. Ghosh and H.-H. S. Lee proposes the technique Smart Refresh \cite{smartrefresh} that targets Access Recency. The basic idea is to maintain a time-out counter for each memory row stored in the MC and these counters is set if the corresponding row is accessed. When a counter time-out, the MC issues a ROR refresh to the corresponding row and also sets the counter.

The counters are decremented in such a fashion so the refreshes are distributed evenly over time. A higher bit size on the counters increases time granularity, which results in that a counter can be set later in time and thus postpones the refresh further.

Can be deactivated

As mentioned, Smart Refresh modifies the MC. The memory footprint of the technique is 0.0024\% of total device capacity when 3-bits counters are used and assumptions , which is stored in the MC.

Results

@@

Implemented so only one line per each logical segments are decremented at the same time, which leads to that the refreshes are evenly distributed over time. (I.e. no burst refresh.)

@@

move:
Smart refresh:
Employed its solution to 2 GB ram. What about todays sizes? How effective is the technique now? - Only works effective if large parts the memory is acccessed, i.e. large part of the memory has to be valid to begin with. Is it so today?
